### 2015-01-10 mcmc.tex
cross-refs
note re Z

In a probabilistic inference, the posterior pdf $p(\pars\given\data)$,
or pdf for the parameters $\pars$ given the data $\data$, is
constructed from the likelihood $p(\data\given\pars)$, or pdf for the
data given the parameters, and the prior pdf $p(\pars)$ for the
parameters by what's often known as ``Bayes rule'',
\begin{eqnarray}
p(\pars\given\data) &=& \frac{1}{Z}\,p(\data\given\pars)\,p(\pars)
\quad .
\end{eqnarray}
In these contexts, the constant $Z$, sometimes written as
$p(\data)$, is known by the names ``evidence'', ``marginal likelihood''
the ``Bayes integral'' and ``prior predictive probability'', and is usually
\emph{extremely hard to
  calculate}.\note{Say words about why we call it $Z$ not $p(\data)$
  and why it is extremely hard to calculate.}
That is, you often know the function $p(\pars\given\data)$ up to a
constant factor; you can compute ratios of the pdf at pairs of points,
but not the precise value at any individual point.

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $d$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf and diagnose problems.

@@

\section{Tuning}

@@

\section{Troubleshooting and advice}

>>>>

In a probabilistic inference, the posterior pdf $p(\pars\given\data)$,
or pdf for the parameters $\pars$ given the data $\data$, is
constructed from the likelihood $p(\data\given\pars)$, or pdf for the
data given the parameters, and the prior pdf $p(\pars)$ for the
parameters by what's often known as ``Bayes rule'',
\begin{eqnarray}
p(\pars\given\data) &=& \frac{1}{Z}\,p(\data\given\pars)\,p(\pars)
\label{eq:bayes}\quad .
\end{eqnarray}
In these contexts, the constant $Z$, sometimes written as
$p(\data)$, is known by the names ``evidence'', ``marginal likelihood''
the ``Bayes integral'' and ``prior predictive probability'', and is usually
\emph{extremely hard to
  calculate}.\note{Some would call
  the factor $Z$ in \equationname~(\ref{eq:bayes}) something like
  ``$p(\data)$'' or ``the marginal likelihood'' or the 
  probability of the data.
  Of course it is only the probability of the data within the model
  space under consideration!
  We discuss this object and these issues [MORE ELSEWHERE; CITE].
  The factor $Z$ is often hugely difficult to compute, because the
  likelihood (or the prior) can have extremely complex structure,
  with multiple arbitrarily compact modes, arbitrarily positioned
  in the (presumably high dimensional) parameter space $\pars$.
  Elsewhere, we discuss the computation of this object (cite Fengji)
  and---perhaps more importantly---why you probably never really
  \emph{want} to compute it.
  The computation of $Z$ is outside the scope of this \documentname,
  but suffice it to say:
  If the computation of $Z$ is in your travel plans, you will need to
  think about implementing some non-trivial integrating or sampling
  technology; we will give some brief description of options in
  \sectionname~\ref{sec:methods}.}
That is, you often know the function $p(\pars\given\data)$ up to a
constant factor; you can compute ratios of the pdf at pairs of points,
but not the precise value at any individual point.

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $d$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf (\sectionname~\ref{sec:tuning})
and diagnose problems (\sectionname~\ref{sec:troubleshooting}).

@@

\section{Tuning}\label{sec:tuning}

@@

\section{Troubleshooting and advice}\label{sec:troubleshooting}
